Motivating Exploration in Reinforcement Learning

### 2020-04-16 paper/main.tex
format figures

\subsection{Implementation Details}
Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \cite{ppo, gym}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments). Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames. We use the Adam optimizer \cite{adam} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project). All of our models are trained synchronously on Nvidia GTX 1080Ti (11GB VRAM) and Tesla V100 (16GB VRAM) gpus with batch size equivalent to the number of parallel environments.

@@

\subsubsection{Dense Extrinsic Reward}
\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
\end{subfigure}
\caption{Here we compare the three A2C variants in the \texttt{Breakout} (left) and \texttt{Seaquest} (right) environments. Rewards are again cumulative over an entire episode.}
\label{fig:seaquest-breakout-reward}
\end{figure}

@@

\subsection{Qualitative Results}
Qualitative Results

For each trained environment, two agents were trained to investigate the effects of exploration. One artificial agent would be trained without exploration and the other would be trained with exploration. Videos of the agents playing the game in the environment were generated, and through these videos, qualitative observations are made with regards to the effects of exploration, and any interesting behaviours that are seen in the agents that leverage exploration.

Qualitative Effects of Exploration

Pong

With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in the figures above. However, the agent with exploration noticeably has a much higher rate of movement compared to the agent without exploration, which could lead to more states of the environment being explored. The agent with exploration often makes many movements that are much faster than the agent without exploration, as can be observed in the supplementary videos.

Breakout

With regards to Breakout, the environment out of the three investigated in this paper with a higher reward density than Pong but a lower reward density than Seaquest, there is a very noticeable difference in the behaviour of the two agents. The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in the figure above. The agent with exploration makes movements all across the screen in order to maximize cumulative reward, and intuitively, explores its environment much more than the agent without exploration. This is a very distinct difference in behaviour that can be clearly observed, especially in the supplementary videos provided.

Seaquest

With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents. This difference between the two agents is the same as the difference noticed in the agents of Breakout. The agent without exploration in Seaquest does not explore its environment at all, and instead opts to stay above sea level at the top left corner. This behaviour is quite unique because it completely avoids any negative rewards in the environment caused by the fish attacking the agent or due to oxygen loss. On the other hand, the agent with exploration makes movements to fully explore the environment to maximize reward while taking the risk of negative rewards as well.

@@

Interesting Behaviours

>>>>

\subsection{Implementation Details}
Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \cite{ppo, gym}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0}, using consistent random seeds for each of the environments. Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames. We use the Adam optimizer \cite{adam} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project). All of our models are trained synchronously on Nvidia GTX 1080Ti (11GB VRAM) and Tesla V100 (16GB VRAM) gpus with batch size equivalent to the number of parallel environments.

@@

\subsubsection{Dense Extrinsic Reward}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
\end{subfigure}
\caption{Here we compare the three A2C variants in the \texttt{Breakout} (left) and \texttt{Seaquest} (right) environments. Rewards are again cumulative over an entire episode.}
\label{fig:seaquest-breakout-reward}
\end{figure}

@@

\subsection{Qualitative Results}
For each trained environment, two agents were considered to investigate the effects of exploration. One artificial agent would be trained without exploration and the other would be trained with exploration. Videos of the agents playing the game in the environment were generated, and through these videos, qualitative observations are made with regards to the effects of exploration, and any interesting behaviours that are seen in the agents that leverage exploration.

\subsubsection{Qualitative Effects of Exploration}

\subsubsubsection{\textbf{Pong}}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/pong}
  \caption{With Exploration}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/pong_no_exp}
  \caption{Without Exploration}
\end{subfigure}
\caption{Pong environment visualizations}
\label{fig:pong}
\end{figure}
With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in Fig \ref{fig:pong}. However, the agent with exploration noticeably has a much higher rate of movement compared to the agent without exploration, which could lead to more states of the environment being explored. The agent with exploration often makes many movements that are much faster than the agent without exploration, as can be observed in the supplementary videos.

\subsubsubsection{\textbf{Breakout}}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/breakout}
  \caption{With Exploration}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/breakout_no_exp}
  \caption{Without Exploration}
\end{subfigure}
\caption{Breakout environment visualizations}
\label{fig:breakout}
\end{figure}
With regards to Breakout, the environment out of the three investigated in this paper with a higher reward density than Pong but a lower reward density than Seaquest, there is a very noticeable difference in the behaviour of the two agents. The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in Fig. \ref{fig:breakout}. The agent with exploration makes movements all across the screen in order to maximize cumulative reward, and intuitively, explores its environment much more than the agent without exploration. This is a very distinct difference in behaviour that can be clearly observed, especially in the supplementary videos provided.

\subsubsubsection{\textbf{Seaquest}}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/seaquest}
  \caption{With Exploration}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.7\linewidth]{figures/seaquest_no_exp}
  \caption{Without Exploration}
\end{subfigure}
\caption{Seaquest environment visualizations}
\label{fig:seaquest}
\end{figure}
With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents as seen in Fig. \ref{fig:seaquest}. This difference between the two agents is the same as the difference noticed in the agents of Breakout. The agent without exploration in Seaquest does not explore its environment at all, and instead opts to stay above sea level at the top left corner. This behaviour is quite unique because it completely avoids any negative rewards in the environment caused by the fish attacking the agent or due to oxygen loss. On the other hand, the agent with exploration makes movements to fully explore the environment to maximize reward while taking the risk of negative rewards as well.

@@

\subsubsection{Interesting Behaviours}
